eda

### 2024-05-15 analysis/eda.Rmd
Next, map it by esac publisher

wos_oct_23_stat <- wos_jct_items |>
  group_by(earlies_year) |>
  summarise(all = n_distinct(item_id),
            core = sum(core_pub),
            hybrid_oa = sum(hybrid_oa),
            cr_oa = sum(cr_oa)) |>
  inner_join(core_hybrid_oa) |>
  mutate(oa_prop = hybrid_oa / all,
         core_oa_prop = core_hybrid_oa / core,
         cr_prop = cr_oa / all,
         core_cr_prop = core_cr_oa / core)

write_csv(wos_oct_23_stat, here::here("data", "wos_oct_23_stat.csv"))
```

>>>>

wos_24_stat <- wos_jct_items |>
  group_by(earlies_year) |>
  summarise(all = n_distinct(item_id),
            core = sum(core_pub),
            hybrid_oa = sum(hybrid_oa),
            cr_oa = sum(cr_oa)) |>
  inner_join(core_hybrid_oa) |>
  mutate(oa_prop = hybrid_oa / all,
         core_oa_prop = core_hybrid_oa / core,
         cr_prop = cr_oa / all,
         core_cr_prop = core_cr_oa / core)

#write_csv(wos_oct_23_stat, here::here("data", "wos_oct_23_stat.csv"))
```

## by journal

```{r}

## Wos data

jn_core_hybrid_oa <- wos_jct_items |>
  filter(core_pub == 1) |>
  group_by(earlies_year, issn_l) |>
  summarise(core_hybrid_oa = sum(hybrid_oa),
            core_cr_oa = sum(cr_oa))

jn_wos_24_stat <- wos_jct_items |>
  group_by(earlies_year, issn_l) |>
  summarise(all = n_distinct(item_id),
            core = sum(core_pub),
            hybrid_oa = sum(hybrid_oa),
            cr_oa = sum(cr_oa)) |>
  inner_join(jn_core_hybrid_oa) |>
  mutate(oa_prop = hybrid_oa / all,
         core_oa_prop = core_hybrid_oa / core,
         cr_prop = cr_oa / all,
         core_cr_prop = core_cr_oa / core)

## hoaddata

cr_all_per_year <- hoaddata::jn_ind |> 
  distinct(issn_l, cr_year, jn_all) |>
  group_by(cr_year, issn_l) |>
  summarise(hoad_articles = sum(jn_all))

cr_oa <- hoaddata::jn_ind |>
  filter(!is.na(cc)) |> 
  group_by(cr_year, issn_l) |>
  summarise(hoad_oa = sum(cc_total))

jn_wos_stats <- inner_join(jn_wos_24_stat, cr_per_year, by = c("earlies_year" = "cr_year", "issn_l"))
```

### Exploration

```{r}
ggplot(jn_wos_stats, aes(hoad_oa / hoad_articles, core_cr_prop)) +
  geom_point(alpha = 0.05) +
  coord_equal() +
  geom_smooth(method = "lm") +
  facet_wrap(~ earlies_year)
```
